---
title: "EPI 560: Lab 4"
author: "Yilda Macias"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  ioslides_presentation:
    incremental: TRUE
    df_print: paged
    theme: cosmo
    toc: yes
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "build") })
---

# EPI 560 Final Project Slides

## Introduction
### Yilda Macias
  * 2nd-year Epi PhD student
  * Dog mom
  * Avid napper
  * Goose supporter

```{r imageintro, echo=FALSE, message=FALSE, warning=FALSE}
library(pacman)
p_load("png")

file_path <- "/Users/yilda/Desktop/EPI_560R/macias_y_epi560/goose.png"
# Read the PNG image
image_data <- readPNG(file_path)
# Display the image
plot(1:2, type = 'n', xlab = "", ylab = "", xlim = c(0, 2.4), ylim = c(0, 1), axes=FALSE)
rasterImage(image_data, 1, 0, 2, 1)
```

## Cars Dataset

```{r codechunk}
print(head(cars))
```

# Markdown Formatting Overview

# Headers

## *You can create different headers by adding a different number of hashtags in the front (#).*

# Header 1
## Header 2
### Header 3
#### Header 4
##### Header 5
###### Header 6

## Lists

### Ordered Lists
  - You can create ordered lists by using either an asterisk or a dash in the beginning. You can create sub-bullets by hitting tab. Here is a to-do list for this weekend with order based on priority level.
      1. 560 Homework
      2. 517 Homework
      3. 516 Mini project
      
### Unordered Lists
  - You can create an unordered list by just including items using bullets. These bullets can be generated using asterisks (*). Here is an example grocery list.
    * Chives
    * Tomatillos
    * Green chile
    * Chicken breast
    * Tortillas
    
### Task Lists
  - Here I added double brackets in the beginning to make this a checkable list. If I include an `x` inside one, when the document is knitted, this will be checked off.
      - [x] **Cry**
      - [ ] *Nap*
      - [ ] *Repeat*
      
     
## Text Formats
  + Get **daisies**
  + Get *tulips*
  + Get $\sf{H_{2}O}$
  + Calculate $\alpha$ and $\gamma$ to complete estimates for `regression equation`: 
  $$\alpha + \beta_1X_1 + \beta_2X_2 + \gamma_1X_1X_2 $$
  
# Embedding images

## **Dexter still smells like beef.**
```{r image, echo=FALSE, message=FALSE, warning=FALSE}
library(pacman)
p_load("png")

file_path1 <- "/Users/yilda/Desktop/EPI_560R/macias_y_epi560/dex.png"
# Read the PNG image
image_data1 <- readPNG(file_path1)
# Display the image
plot(1:2, type = 'n', xlab = "", ylab = "", xlim = c(0, 2.4), ylim = c(0, 1), axes=FALSE)
rasterImage(image_data1, 0, 0, 1, 1)
```

# Code Chunks
```{r library, echo=FALSE, message=FALSE, warning=FALSE}
# Setup, load libraries needed
library(pacman)
p_load("tidyverse", "knitr", "ggplot2")


# Generate variables for dataframe
## set seed
set.seed(35)

## generate exposure, outcome, and confounder
exposure <- sample(c(TRUE, FALSE), size = 20, replace = TRUE, prob = c(0.4, 0.6))
outcome <- sample(c(TRUE, FALSE), size = 20, replace = TRUE, prob = c(0.7, 0.3))
confounder <- round(ifelse(exposure, rnorm(sum(exposure), mean = 5, sd = 2), rnorm(20 - sum(exposure), mean = 8, sd = 2)),2)

## create dataframe
data <- data.frame(
  exposure = exposure,
  outcome = outcome,
  confounder = confounder
)

# view(data)
```

## Formatted table of the first 5 rows of your small dataframe using the kable function
```{r table5, echo=FALSE, message=FALSE, warning=FALSE}
# Print 5 rows of data
kable(head(data, 5))
```

## 2x2 table showing exposure and outcome
```{r twobtwo, echo=FALSE, message=FALSE, warning=FALSE}
# Create a 2x2 contingency table
data2 <- table(data$exposure, data$outcome)
data2 <- as.data.frame.matrix(data2)

# Add titles to rows and columns
rownames(data2) <- c("Unexposed", "Exposed")
colnames(data2) <- c("No Outcome", "Outcome") 

# Display table
kable(data2, caption = "2 by 2 Table")
```

## Histogram of confounder
```{r hist1, echo=FALSE, message=FALSE, warning=FALSE}
# Print histogram of confounder
ggplot(data, aes(x = confounder)) +
  geom_histogram(binwidth = 1, fill = "darkred", color="black") +
  labs(title = "Histogram of Confounder", x = "Confounder Value", y = "Frequency (counts)")
```

## Histogram of confounder by exposure
```{r hist2, echo=FALSE, message=FALSE, warning=FALSE}
# Stratify by exposure
ggplot(data, aes(x = confounder)) +
  geom_histogram(binwidth = 1, fill = "darkred", color="black") +
  labs(title = "Histogram of Confounder", x = "Confounder Value", y = "Frequency (counts)") +
  facet_grid(. ~ exposure, labeller = labeller(exposure = c("TRUE" = "Exposed", "FALSE" = "Unexposed")))
```

## Inline Code 
In the dataset generated above, the `confounder` has a mean value of `r mean(data$confounder[data$exposure == TRUE])` in the exposed, and a `r mean(data$confounder[data$exposure == FALSE])` in the unexposed. Although the instructions said that the distributions for the confounder in both exposed and unexposed was meant to be normally distributed, I was not able to figure this out.


\pagebreak  

## Code Appendix  
 
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE} 
```
