---
title: "EPI 560: Lab 7"
author: "Yilda Macias"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
        df_print: paged
        theme: cosmo
        toc: yes
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "build") })
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Setup
library(pacman)
p_load("tidyverse", "knitr", "dplyr", "ggplot2", "haven", "readxl", "rhdf5", "arrow")

# if (!require("BiocManager", quietly = TRUE))
# install.packages("BiocManager")
# BiocManager::install("rhdf5")

# install.packages("arrow", repos = c("https://apache.r-universe.dev"))
```

## Lab 7: Responses 
### Part 1.
All the data appear to be in each of datasets created from reading in the different formats.

```{r csv, echo=FALSE, message=FALSE, warning=FALSE}
# code for part 1

# load csv
before.csv <- Sys.time()
csv <- read.csv("P_DEMO.csv")
after.csv <- Sys.time()
(elapsed.csv <- after.csv - before.csv)

# scatterplot
csvplot <- ggplot(csv, aes(x=WTINTPRP, y=WTMECPRP)) +
  geom_point() +
  labs(title = "Sample weights used in NHANES (csv)", x = "WTINTPRP", y = "WTMECPRP")
csvplot
```


```{r zip, echo=FALSE, message=FALSE, warning=FALSE}
# load zip
before.zip <- Sys.time()
zipp <- read.csv(unz("/Users/yilda/Desktop/EPI_560R/macias_y_epi560/P_DEMO.csv.zip", "P_DEMO.csv"))
after.zip <- Sys.time()
(elapsed.zip <- after.zip - before.zip)

# scatterplot
zipplot <- ggplot(zipp, aes(x=WTINTPRP, y=WTMECPRP)) +
  geom_point() +
  labs(title = "Sample weights used in NHANES (zip)", x = "WTINTPRP", y = "WTMECPRP")
zipplot
```

```{r dta, echo=FALSE, message=FALSE, warning=FALSE}
# load dta
before.dta <- Sys.time()
dta <- read_dta("P_DEMO.dta")
after.dta <- Sys.time()
(elapsed.dta <- after.dta - before.dta)

# scatterplot
dtaplot <- ggplot(dta, aes(x=WTINTPRP, y=WTMECPRP)) +
  geom_point() +
  labs(title = "Sample weights used in NHANES (dta)", x = "WTINTPRP", y = "WTMECPRP")
dtaplot
```


```{r feather, echo=FALSE, message=FALSE, warning=FALSE}
# load feather
before.feather <- Sys.time()
feather <- arrow::read_feather("P_DEMO.feather")
after.feather <- Sys.time()
(elapsed.feather <- after.feather - before.feather)

# scatterplot
featherplot <- ggplot(feather, aes(x=WTINTPRP, y=WTMECPRP)) +
  geom_point() +
  labs(title = "Sample weights used in NHANES (feather)", x = "WTINTPRP", y = "WTMECPRP")
featherplot
```

```{r hdf5, echo=FALSE, message=FALSE, warning=FALSE}
# load hdf5
before.hdf5 <- Sys.time()
hdf5 <- rhdf5::h5read("P_DEMO.hdf5", name="P_DEMO")
after.hdf5 <- Sys.time()
(elapsed.hdf5 <- after.hdf5 - before.hdf5)

# scatterplot
hdf5plot <- ggplot(hdf5, aes(x=WTINTPRP, y=WTMECPRP)) +
  geom_point() +
  labs(title = "Sample weights used in NHANES (hdf5)", x = "WTINTPRP", y = "WTMECPRP")
hdf5plot
```

```{r parquet, echo=FALSE, message=FALSE, warning=FALSE}
# load parquet
before.parquet <- Sys.time()
parquet <- arrow::read_parquet("P_DEMO.parquet", name="P_DEMO")
after.parquet <- Sys.time()
(elapsed.parquet <- after.parquet - before.parquet)

# scatterplot
parquetplot <- ggplot(parquet, aes(x=WTINTPRP, y=WTMECPRP)) +
  geom_point() +
  labs(title = "Sample weights used in NHANES (parquet)", x = "WTINTPRP", y = "WTMECPRP")
parquetplot
```

```{r xlsx, echo=FALSE, message=FALSE, warning=FALSE}
# load xlsx
before.xlsx <- Sys.time()
xlsx <- read_xlsx("P_DEMO.xlsx")
after.xlsx <- Sys.time()
(elapsed.xlsx <- after.xlsx - before.xlsx)

# scatterplot
xlsxplot <- ggplot(xlsx, aes(x=WTINTPRP, y=WTMECPRP)) +
  geom_point() +
  labs(title = "Sample weights used in NHANES (xslx)", x = "WTINTPRP", y = "WTMECPRP")
xlsxplot
```

### Part 2. 
The file type that read in the fastest was feather, at 0.028 seconds. It was followed by the parquet file, at 0.032 seconds, for a difference of `r round(elapsed.parquet-elapsed.feather,2)`. I tried using inline code for this, but it did not work. I am not sure why. Is it because the ouput is not a simple estimate? The real answer should be 0.004 seconds.


\pagebreak  

## Code Appendix  
 
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE} 
```
